Robust Control of a Fluxonium Qubit

### 2020-11-25 paper/0abstract.tex
added lots of references to introduction

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zachary Manchester}
\affiliation{
  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
  % Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high-fidelity gates on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control techniques have proven effective in experimentally
  realizing high-fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  parameter deviations and noise.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentiation.
  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
  integrating the Lindblad master equation, which is computationally expensive.
  We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We demonstrate these techniques numerically on a fluxonium qubit with realistic
  experimental parameters and constraints,
  achieving orders of magnitude gate error reductions from our baseline gate set.
\end{abstract}

>>>>

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zachary Manchester}
\affiliation{
  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  The ability to engineer high-fidelity gates on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control techniques have proven effective in experimentally
  realizing high-fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  parameter deviations and noise.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentiation.
  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
  integrating the Lindblad master equation, which is computationally expensive.
  We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We demonstrate these techniques numerically on a fluxonium qubit with realistic
  experimental parameters and constraints,
  achieving orders of magnitude gate error reductions from our baseline gate set.
\end{abstract}

### 2020-11-25 paper/1introduction.tex
added lots of references to introduction

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
For quantum computation
QOC techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system's trajectory to deviate from that predicted in
optimization, hampering performance.
Robust control techniques 
address this issue by encoding
experimental errors in optimization objectives, yielding performance
garauntees in the presence of model uncertainty \cite{Zhou97,Morimoto00,Manchester18}.
In this work we employ robust control techniques to mitigate
realistic system parameter deviations and decoherence that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates \todo{citations needed}. 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
In order to address these issues,
recent work has sought to mitigate
errors due to parameter deviations.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
longitudinal relaxation and pure dephasing decoherence
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work, we study three robust control techniques that
make the optimized trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
\item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling,
  rembold2020introduction}.
  \item An unscented sampling method, adapted from the unscented tranform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
  to compute derivative information of parameter deviations with respect to the 
  quantum state trajectory.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

>>>>

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
\todo{more nmr and sqc cite} Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{
  brouzos2015quantum, de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
  treutlein2006microwave, van2016optimal}, nitrogen-vacancy centers in
diamond \cite{chou2015optimal, dolde2014high, geng2016experimental,
  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
For quantum computation,
optimal control techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system to deviate
from the model used in optimization, leading
to poor experimental performance.
Robust control techniques improve upon
standard optimal control techniques by encoding
model uncertainties
in optimization objectives, yielding performance
guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
We adapt robust control techniques from the robotics community to mitigate
realistic model uncertainties for
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have had tremendous success in
engineering high fidelity quantum gates \cite{
  chakram2020multimode, chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
  gokhale2019partial, grace2007optimal, heeres2017implementing, huang2014optimal,
  kelly2014optimal, leng2019robust, liebermann2016optimal,
  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
  spiteri2018quantum, sporl2007optimal}. 
While traditional QOC techniques can predict system behavior
with high accuracy, they are sensitive to experimental
errors such as parameter drift, noise, finite control 
resolution, and decoherence. Analytic techniques to mitigate parameter deviation errors
include dynamic and geometric phase considerations
\cite{han2020experimental, xu2020nonadiabatic} and
composite pulses \cite{merrill2014progress}.
To mitigate decoherence, Floquet techniques have been employed
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques to mitigate parameter deviation errors
include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
  wittler2020integrated} and open-loop methods \cite{
  allen2019robust, carvalho2020error, reinhold2019controlling,
  rembold2020introduction, kosut2013robust}.
Numerical techniques to mitigate decoherence include
modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.

In this work, we study three robust control techniques that
make the system's state trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{allen2019robust,
    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
    to compute derivative information of parameter deviations with respect to the 
    quantum state trajectory.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular, taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation type decoherence and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.
